---
title: "A deep dive"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width=4, fig.height=3) 
library(dplyr)
library(ggplot2)

community <- "bbs_rtrg_1_4"

ibd_filepath <- paste0("C:\\Users\\diaz.renata\\Documents\\GitHub\\BBSsize\\analysis\\isd_data\\ibd_isd_", community, ".Rds")

ibd <- readRDS(ibd_filepath)

```

Here I have loaded the IBD which includes a size estimate and metabolic rate estimate, and species ID, for every individual at every time point. 

```{r head ibd}
head(ibd) 
```

We can use the IBD to compute aggregate metrics per year.

```{r aggregates}

agg <- ibd %>%
  group_by(year) %>%
  summarize(totalind = dplyr::n(),
            nspp = length(unique(id)),
            totalmass = sum(ind_size),
            totalenergy = sum(ind_b),
            meanmass = mean(ind_size),
            meanenergy = mean(ind_b)) %>%
  ungroup()

head(agg)

```

Let's plot the aggregate metrics over time.

```{r agg over time, fig.dim = c(4,3)}
ggplot(agg, aes(year, nspp)) +
  geom_point(color = "pink") +
    geom_line(color = "pink") +
  geom_smooth(method = "lm", se = F, color = "pink") +
  theme_bw() +
  ggtitle("Species richness")


ggplot(agg, aes(year, totalind)) +
  geom_point(color = "blue") +
    geom_line(color = "blue") +
    geom_smooth(method = "lm", se = F, color = "blue") +
  theme_bw() +
  ggtitle("Total individual abundance")


ggplot(agg, aes(year, totalmass)) +
  geom_point(color = "purple") +
    geom_line(color = "purple") +
    geom_smooth(method = "lm", se = F, color = "purple") +
  theme_bw() +
  ggtitle("Total biomass")


ggplot(agg, aes(year, totalenergy)) +
  geom_point(color = "red") +
    geom_line(color = "red") +
    geom_smooth(method = "lm", se = F, color = "red") +
  theme_bw() +
  ggtitle("Total energy")

ggplot(agg, aes(totalind, totalenergy, color = totalmass)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c()

ggplot(agg, aes(year, meanmass)) +
  geom_point(color = "purple") +
    geom_line(color = "purple") +
    geom_smooth(method = "lm", se = F, color = "purple") +
  theme_bw() +
  ggtitle("Mean biomass")


ggplot(agg, aes(year, meanenergy)) +
  geom_point(color = "red") +
    geom_line(color = "red") +
    geom_smooth(method = "lm", se = F, color = "red") +
  theme_bw() +
  ggtitle("Mean energy")
```

```{r agg lms}

agg_for_lm <- agg %>%
  select(year, totalind, totalmass, totalenergy) %>%
  tidyr::pivot_longer(-year, names_to = "currency") %>%
  group_by(currency) %>%
  mutate(value = scale(value)) %>%
  ungroup()

ggplot(agg_for_lm, aes(x = year, y = value, color = currency)) +
  geom_point() +
  geom_line() +
  theme_bw()

currency_lm <- lm(value ~ year * currency, data = agg_for_lm)

summary(currency_lm)

plot(currency_lm)

```


The story I would tell from this is:

- Overall declines in abundance more or less matching individuals and energy metrics; not a strong systematic functional shift. 




Let's look at the ISD by year? Will be big!!

```{r isds, fig.dim = c(20,20)}

ggplot(ibd, aes(x = log(ind_size))) +
  geom_density() +
  theme_bw() +
  facet_wrap(vars(year), scales = "free_y", ncol = 4)

```

Just based on intuition, let's contrast this with species histograms by year...

```{r composition, fig.dim = c(20,20)}

ggplot(ibd, aes(x = id, y = ..count..)) +
  geom_bar() +
  theme_bw() +
  facet_wrap(vars(year), scales = "free_y", ncol = 4)

```

So now I would like to see plots of:

* Pairwise ISD overlap (y) against distance in time (x). 
* Euclidean community distance (y) against distance in time (x).

It looks to me almost like the ISDs could be plausibly unordered: not shifting directionally through time. Such that overlap should generally be high, and there should not be a relationship between how far apart years are and how similar they are. But that we might have seen directional change in species composition, meaning that years farther apart will be less similar than adjacent ones.

This isn't what we get from these plots! Not sure if that's cause I'm wrong, or because these plots aren't sophisticated enough to handle autocorrelation etc.

The one piece of this I think is really interesting (and I have some confidence in) is, that Euclidean distance and overlap appear pretty uncorrelated. I *think* this points to a large degree of functional replacement of species: you can have high overlap with large distance, if species 1 has replaced species 2 but they are the same size.

```{r isd overlap pairwise, fig.dim = c(4,3)}
library(isds)

years_key <- data.frame(year = unique(agg$year),
                        species = 1:length(unique(agg$year)))

all_pairings <- get_pairs(nrow(years_key))

ibd_for_pairings <- ibd %>%
  left_join(years_key) %>%
  rename(wgt = ind_size) 

all_species_pairs <- apply(all_pairings, MARGIN = 1, FUN = get_species_pairs, community_df = ibd_for_pairings)

all_pair_overlaps <- lapply(all_species_pairs, FUN = isds:::pair_overlap, min_size = 0, max_size = max(ibd_for_pairings$wgt) * 1.5, npoints = 1024)

all_pairings$overlap <- unlist(all_pair_overlaps)

all_pairings$distance <- NA


community_matrix <- ibd %>%
  group_by(year, id) %>%
  summarize(abund = (dplyr::n())) %>%
  ungroup() %>%
  tidyr::pivot_wider(names_from = id, values_from = abund, values_fill = 0) %>%
  select(-year)

community_matrix <- as.matrix(community_matrix)

for(i in 1:nrow(all_pairings)) {
  
  this_pair <- community_matrix[ c(all_pairings$sp1[i], all_pairings$sp2[i]),]
  
  all_pairings$distance[i] <- dist(this_pair)[[1]]

}

ibd_overlap <- all_pairings %>%
  rename(species = sp1) %>%
  left_join(years_key) %>%
  rename(year1 = year,
         sp1 = species) %>%
  rename(species = sp2) %>%
  left_join(years_key) %>%
  rename(year2 = year, 
         sp2 = species)


ibd_overlap <- ibd_overlap %>%
  mutate(year_distance = abs(year1 - year2))
```


```{r overlap v eucl}
ggplot(ibd_overlap, aes(distance, overlap)) +
  geom_point() +
  theme_bw()
```

Similarity in species composition doesn't appear to predict ISD overlap.


### Scrap
```{r year dist v overlap}
ggplot(ibd_overlap, aes(year_distance, overlap)) + 
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm", se = F)

```

OK, so it looks like there's a weak decrease in overlap for years that are farther apart.

```{r distance v eucl}

ggplot(ibd_overlap, aes(year_distance, distance)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm", se = F)

```

And if anything, years far apart are *closer* in Euclidean space.

```{r since 1972 overlap}
ggplot(filter(ibd_overlap, year1 == 1972), aes(year_distance, overlap)) + 
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm", se = F) +
  ggtitle("Time since 1972")
```

Looks like a drift away from starting conditions over time, but note huge noise.

```{r eucl since 1972}
ggplot(filter(ibd_overlap, year1 == 1972), aes(year_distance, distance)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm", se = F)+
  ggtitle("Time since 1972")
```

In contrast composition doesn't obviously drift, but still extremely noisy.

```{r consecutive overlap}
consecutive_years <- years_key %>%
  select(-species) %>%
  rename(year1 = year)

consecutive_years$year2 <- c(consecutive_years$year[2:17], NA)


consecutive <- ibd_overlap %>%
  right_join(consecutive_years)
  
ggplot(consecutive, aes(x = year1, y = overlap)) +
  geom_point()  +
  theme_bw() +
  geom_smooth(method = "lm", se = F)
```

Consecutive samples overlap *less* over time wrt ISD.

```{r consecutive eucl}
ggplot(consecutive, aes(x = year1, y = distance)) +
  geom_point()  +
  theme_bw() +
  geom_smooth(method = "lm", se = F)

```

Consecutive samples are *more* similar in species composition over time.
