---
title: "a trimmed-ish draft"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.dim = c(5,3))

library(dplyr)
library(gratia)
library(ggplot2)
load_mgcv()

ts <- read.csv(here::here("gams", "working_datasets.csv"))

unique_sites <- unique(ts$site_name)

site_dfs <- lapply(unique_sites, FUN = function(site, full_ts) return(filter(full_ts, site_name == site)), full_ts = ts)

source(here::here("gams", "gam_fxns", "wrapper_fxns.R"))
source(here::here("gams", "gam_fxns", "sunrise_fxns.R"))

```

#### With portal

```{r just e portal, echo = F}

rats <- filter(ts, site_name == "portal_rats")

portal_mean_perc_e <- sum(rats$energy) / sum(rats$abundance)

rats <- rats %>% 
  mutate(scaled_energy = energy / portal_mean_perc_e)

rats_long <- rats %>%
  select(year, energy, abundance, scaled_energy) %>%
  tidyr::pivot_longer(-year, names_to = "currency", values_to = "value")

ggplot(filter(rats_long, currency != "energy"), aes(year, value, color = currency)) +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d()

```


```{r}
e_mod <- mod_wrapper(rats, response_variable = "energy", identifier = "site_name", k = 5)
re_e_mod <- mod_wrapper(rats, response_variable = "scaled_energy", identifier = "site_name", k = 5)
n_mod <- mod_wrapper(rats, response_variable = "abundance", identifier = "site_name", k = 5)

e_samples <- samples_wrapper(re_e_mod, seed_seed = 1994)
n_samples <- samples_wrapper(n_mod, seed_seed = 1990)

joint_samples <- bind_rows(e_samples, n_samples) %>%
  select(year, currency, mean, upper, lower) %>%
  distinct()

ggplot(joint_samples, aes(year, mean, color = currency, fill = currency)) +
  geom_line(aes(year, mean)) +
  geom_ribbon(aes(year, ymin = lower, ymax = upper),  alpha = .25) +
  ggtitle("Fits for energy and abundance") +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d()

```


Calculated off the samples, we can get:

* the net change from beginning to end, plus a CI
* a timeseries of the instantaneous rate of change, which we can summarize in order to get a sense of how much backtracking is occurring over the course of the timeseries

## Net change

```{r shortcut to net change}


e_change <- net_change_wrapper(e_samples)
n_change <- net_change_wrapper(n_samples)

both_change <- bind_rows(e_change, n_change)

ggplot(both_change, aes(currency, net_proportional)) +
  geom_boxplot() +
  theme_bw()

both_summary <- change_summary_wrapper(both_change)

both_summary
```

So this is a way of saying

* Abundance increases by about ```r both_summary$mean_net_proportional[1] * 100``` percent, and energy increases by about ```r both_summary$mean_net_proportional[2] * 100```, with a CI given by the lower/upper.

## Instantaneous change and backtracking

```{r instant change}


e_instant_change <- instant_change_wrapper(e_samples)
n_instant_change <- instant_change_wrapper(n_samples)

both_instant_change <- bind_rows(e_instant_change, n_instant_change)

both_ichange_summary <- instant_change_summary_wrapper(both_instant_change)

ggplot(both_ichange_summary, aes(year, mean_ichange_proportional, color = currency, fill = currency)) +
  geom_line(aes(year, mean_ichange_proportional)) +
  geom_ribbon(aes(year, ymin = lower_ichange_proportional, ymax = upper_ichange_proportional),  alpha = .25) +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  geom_hline(yintercept = 0)

both_ichange_ts_summary <- instant_change_absolute_summary_wrapper(both_instant_change)


ggplot(both_ichange_ts_summary, aes(currency, mean_abs_ichange_over_ts)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  theme_bw()
```

On average there's about a 12% change (up or down) timestep to timestep for abundance, and maybe a 9% change for energy. For abundance this is consistently an increase; for energy this is sometimes a decrease and later an increase. 


# For all sites

```{r}

re_e_mods <- list()
n_mods <- list()
e_samples <- list()
n_samples <- list()
joint_samples <- list()
both_change <- list()
both_summary <- list()
both_instant_change <- list()
both_ichange_summary <- list()
both_ichange_ts_summary <- list()

for(i in 1:length(site_dfs)) {
  this_df <- site_dfs[[i]]
  
  
mean_perc_e <- sum(this_df$energy) / sum(this_df$abundance)

this_df <- this_df %>% 
  mutate(scaled_energy = energy / mean_perc_e)

re_e_mods[[i]] <- mod_wrapper(this_df, response_variable = "scaled_energy", identifier = "site_name", k = 5)
n_mods[[i]] <- mod_wrapper(this_df, response_variable = "abundance", identifier = "site_name", k = 5)

e_samples[[i]] <- samples_wrapper(re_e_mods[[i]], seed_seed = 1994)
n_samples[[i]] <- samples_wrapper(n_mods[[i]], seed_seed = 1990)

joint_samples[[i]] <- bind_rows(e_samples, n_samples) %>%
  select(year, currency, mean, upper, lower, identifier) %>%
  distinct()


e_change <- net_change_wrapper(e_samples[[i]])
n_change <- net_change_wrapper(n_samples[[i]])

both_change[[i]] <- bind_rows(e_change, n_change)

both_summary[[i]] <- change_summary_wrapper(both_change[[i]])


e_instant_change <- instant_change_wrapper(e_samples[[i]])
n_instant_change <- instant_change_wrapper(n_samples[[i]])

both_instant_change[[i]] <- bind_rows(e_instant_change, n_instant_change)

both_ichange_summary[[i]] <- instant_change_summary_wrapper(both_instant_change[[i]])

both_ichange_ts_summary[[i]] <- instant_change_absolute_summary_wrapper(both_instant_change[[i]])
}




```

```{r}

joint_samples <- bind_rows(joint_samples)

ggplot(joint_samples, aes(year, mean, color = currency, fill = currency)) +
  geom_line(aes(year, mean)) +
  geom_ribbon(aes(year, ymin = lower, ymax = upper),  alpha = .25) +
  ggtitle("Fits for energy and abundance") +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() + 
  facet_wrap(vars(identifier), scales = "free")

both_change <- bind_rows(both_change)
# 
# ggplot(both_change, aes(identifier, net_proportional, color = currency)) +
#   geom_boxplot() +
#   theme_bw()

both_summary <- bind_rows(both_summary) 

both_summary_wide <- both_summary %>%
  tidyr::pivot_wider(id_cols = identifier, names_from = currency, values_from = c(mean_net_proportional, lower_net_proportional, upper_net_proportional))

ggplot(both_summary_wide, aes(x = mean_net_proportional_abundance, y = mean_net_proportional_scaled_energy, color = identifier)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_net_proportional_scaled_energy,
                    ymax = upper_net_proportional_scaled_energy)) +
  geom_errorbarh(aes(xmin = lower_net_proportional_abundance,
                     xmax = upper_net_proportional_abundance)) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) + 
  geom_abline(slope = 1, intercept = 0)

```

```{r}
both_ichange_summary <- bind_rows(both_ichange_summary)
ggplot(both_ichange_summary, aes(year, mean_ichange_proportional, color = currency, fill = currency)) +
  geom_line(aes(year, mean_ichange_proportional)) +
  geom_ribbon(aes(year, ymin = lower_ichange_proportional, ymax = upper_ichange_proportional),  alpha = .25) +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(identifier), scales = "free")


both_ichange_ts_summary <- bind_rows(both_ichange_ts_summary)

ggplot(both_ichange_ts_summary, aes(currency, mean_abs_ichange_over_ts)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  theme_bw() +
  facet_wrap(vars(identifier))

ggplot(both_ichange_ts_summary, aes(identifier, mean_abs_ichange_over_ts)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(vars(currency))
```

# For all sites, k = 10

```{r}

re_e_mods <- list()
n_mods <- list()
e_samples <- list()
n_samples <- list()
joint_samples <- list()
both_change <- list()
both_summary <- list()
both_instant_change <- list()
both_ichange_summary <- list()
both_ichange_ts_summary <- list()

for(i in 1:length(site_dfs)) {
  this_df <- site_dfs[[i]]
  
  
mean_perc_e <- sum(this_df$energy) / sum(this_df$abundance)

this_df <- this_df %>% 
  mutate(scaled_energy = energy / mean_perc_e)

re_e_mods[[i]] <- mod_wrapper(this_df, response_variable = "scaled_energy", identifier = "site_name", k = 10)
n_mods[[i]] <- mod_wrapper(this_df, response_variable = "abundance", identifier = "site_name", k = 10)

e_samples[[i]] <- samples_wrapper(re_e_mods[[i]], seed_seed = 1994)
n_samples[[i]] <- samples_wrapper(n_mods[[i]], seed_seed = 1990)

joint_samples[[i]] <- bind_rows(e_samples, n_samples) %>%
  select(year, currency, mean, upper, lower, identifier) %>%
  distinct()


e_change <- net_change_wrapper(e_samples[[i]])
n_change <- net_change_wrapper(n_samples[[i]])

both_change[[i]] <- bind_rows(e_change, n_change)

both_summary[[i]] <- change_summary_wrapper(both_change[[i]])


e_instant_change <- instant_change_wrapper(e_samples[[i]])
n_instant_change <- instant_change_wrapper(n_samples[[i]])

both_instant_change[[i]] <- bind_rows(e_instant_change, n_instant_change)

both_ichange_summary[[i]] <- instant_change_summary_wrapper(both_instant_change[[i]])

both_ichange_ts_summary[[i]] <- instant_change_absolute_summary_wrapper(both_instant_change[[i]])
}




```

```{r}

joint_samples <- bind_rows(joint_samples)

ggplot(joint_samples, aes(year, mean, color = currency, fill = currency)) +
  geom_line(aes(year, mean)) +
  geom_ribbon(aes(year, ymin = lower, ymax = upper),  alpha = .25) +
  ggtitle("Fits for energy and abundance") +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() + 
  facet_wrap(vars(identifier), scales = "free")

both_change <- bind_rows(both_change)
# 
# ggplot(both_change, aes(identifier, net_proportional, color = currency)) +
#   geom_boxplot() +
#   theme_bw()

both_summary <- bind_rows(both_summary) 

both_summary_wide <- both_summary %>%
  tidyr::pivot_wider(id_cols = identifier, names_from = currency, values_from = c(mean_net_proportional, lower_net_proportional, upper_net_proportional))

ggplot(both_summary_wide, aes(x = mean_net_proportional_abundance, y = mean_net_proportional_scaled_energy, color = identifier)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_net_proportional_scaled_energy,
                    ymax = upper_net_proportional_scaled_energy)) +
  geom_errorbarh(aes(xmin = lower_net_proportional_abundance,
                     xmax = upper_net_proportional_abundance)) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) + 
  geom_abline(slope = 1, intercept = 0)

```

```{r}
both_ichange_summary <- bind_rows(both_ichange_summary)
ggplot(both_ichange_summary, aes(year, mean_ichange_proportional, color = currency, fill = currency)) +
  geom_line(aes(year, mean_ichange_proportional)) +
  geom_ribbon(aes(year, ymin = lower_ichange_proportional, ymax = upper_ichange_proportional),  alpha = .25) +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(identifier), scales = "free")


both_ichange_ts_summary <- bind_rows(both_ichange_ts_summary)

ggplot(both_ichange_ts_summary, aes(currency, mean_abs_ichange_over_ts)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  theme_bw() +
  facet_wrap(vars(identifier))

ggplot(both_ichange_ts_summary, aes(identifier, mean_abs_ichange_over_ts)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(vars(currency))
```
