---
title: "BBS route that goes by Hartland"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(bbstrends)
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
source(here::here("analysis", "demo_fxns.R"))
```


### Load specific route

The New Hartford route goes up and down Riverton Road and was started in 1994. It feels pretty auspicious. It is route 102, region 18.

```{r load route}

ibd <- readRDS("~/Documents/GitHub/BBSsize/analysis/isd_data/ibd_isd_bbs_rtrg_102_18.Rds")

#ibd <- read.csv("~/Documents/GitHub/BBSsize/analysis/isd_data/isd_bbs_rtrg_7_4.csv")

ibd <- ibd %>%
  mutate(ind_size = abs(ind_size)) %>%
  mutate(log_size = log(ind_size))

```

Here are the species present in this route over the past 25 years:

```{r species list for fun}

ibd_species <- ibd %>%
  select(id, ind_size) %>%
  group_by(id) %>%
  summarize(mean_size = mean(ind_size)) %>%
  ungroup() %>%
  arrange(mean_size)

```


Here is how species richness, abundance, biomass, and energy have changed over those years:

```{r state variables}
# 
# sv <- ibd %>%
#   group_by(year) %>%
#   summarize(abundance = dplyr::n(),
#             biomass = sum(ind_size),
#             energy = sum(ind_b)) %>%
#   ungroup() %>%
#   mutate(         mean_energy = energy/abundance)
# 

sv <- ibd %>%
  group_by(year) %>%
  summarize(abundance = dplyr::n(),
            biomass = sum(ind_size)) %>%
  ungroup() %>%
  mutate(         mean_size = biomass/abundance)


sv_long <- sv %>%
  tidyr::pivot_longer(-year, names_to = "currency", values_to = "value")

ggplot(filter(sv_long), aes(x = year, y = value, color = currency)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d(end = .8) +
  theme(legend.position = "none") +
  facet_wrap(vars(currency), scales = "free_y", ncol = 1)

```

## ISD 

Using the logarithm of mass. 

```{r isd}

ggplot(ibd, aes(log_size)) +
  geom_density()

first5 <- unique(ibd$year)[1:5]
last5 <- sort(unique(ibd$year), decreasing = T)[1:5]

first5_isd <- ibd %>%
  filter(year %in% first5) %>%
  mutate(period = "first5")

last5_isd <- ibd %>%
  filter(year %in% last5) %>%
  mutate(period = "last5")

isds <- bind_rows(first5_isd, last5_isd)

ggplot(isds, aes(log_size, color = year, group = year)) +
  geom_density() +
  facet_wrap(vars(period), scales = "free_y")

min_size = 0
max_size = max(isds$log_size) + .5
npoints = 2000

smooth_first5_isd <- data.frame(log_size = seq(min_size, max_size, length.out = npoints),
                                period = "first5") %>%
  mutate(density = make_kde(first5_isd$log_size, min_size ,max_size, npoints))


smooth_last5_isd <- data.frame(log_size = seq(min_size, max_size, length.out = npoints),
                                period = "last5") %>%
  mutate(density = make_kde(last5_isd$log_size, min_size ,max_size, npoints))


smooth_isds <- bind_rows(smooth_first5_isd, smooth_last5_isd)

ggplot(smooth_isds, aes(log_size, density, color = period)) +
  geom_line() 

overlap <- pair_overlap(list(sp1=first5_isd$log_size, sp2=last5_isd$log_size), min_size, max_size, npoints)
```

The overlap for these is ```r overlap```.

Resampling from the combined KDE:

```{r}

overall_isd <- data.frame(log_size = seq(min_size, max_size, length.out = npoints),
                          density = make_kde(isds$log_size, min_size ,max_size, npoints))

sum(overall_isd$density)

ggplot(overall_isd, aes(log_size, density)) +
  geom_line()
# 
simulated_first5 = replicate(100, expr = sample(overall_isd$log_size, size = nrow(first5_isd), prob = overall_isd$density, replace = T))

simulated_last5 = replicate(100, expr = sample(overall_isd$log_size, size = nrow(last5_isd), prob = overall_isd$density, replace = T))
# # 
# simulated_first5 = replicate(100, expr = sample(c(last5_isd$log_size, first5_isd$log_size), size = nrow(first5_isd), replace = T))
# 
# simulated_last5 = replicate(100, expr = sample(c(last5_isd$log_size, first5_isd$log_size), size = nrow(last5_isd), replace = T))


simulated_overlaps <- vector()

for(i in 1:100) {
  
  simulated_overlaps[i] <- pair_overlap(list(sp1=simulated_first5[,i], sp2=simulated_last5[,i]), min_size, max_size, npoints)
  
}

simulated_kdes <- bind_rows(
  mutate(as.data.frame(apply(simulated_first5, MARGIN =2, FUN = make_kde, min_size = min_size, max_size = max_size, npoints = npoints)), period = "first5", log_size = seq(min_size, max_size, length.out = npoints)),
  mutate(as.data.frame(apply(simulated_last5, MARGIN =2, FUN = make_kde, min_size = min_size, max_size = max_size, npoints = npoints)), period = "last5",log_size = seq(min_size, max_size, length.out = npoints)))
simulated_kdes <- simulated_kdes %>%
  tidyr::pivot_longer(1:100, names_to = "sim", values_to = "density") %>%
  mutate(period_sim = paste0(period, sim))

ggplot(filter(simulated_kdes, sim %in% unique(simulated_kdes$sim)[1:10]), aes(log_size, density, color = period, group = period_sim)) + 
  geom_line(alpha = .5) +
  scale_color_viridis_d(end = .8) +
  geom_line(data = smooth_isds, aes(log_size, density, group= period, linetype = period), inherit.aes = F)


overlaps <- data.frame(source = c(rep("sim", times = 100), "actual"),
                       overlap = c(simulated_overlaps, overlap))
ggplot(filter(overlaps, source == "sim"), aes(overlap)) +
  geom_density() +
  geom_vline(xintercept = filter(overlaps, source == "actual")$overlap, color = "red")

```
### State variable change first-last

```{r}

sv_periods <- data.frame(first5 = first5,
                         last5 = last5) %>%
  tidyr::pivot_longer(everything(), names_to = "period", values_to = "year")

sv_ends_long <- filter(sv_long, year %in% sv_periods$year) %>%
  right_join(sv_periods) 
sv_ends <- filter(sv, year %in% sv_periods$year) %>%
  right_join(sv_periods) %>%
  mutate(period = as.factor(period))


wilcox.test(abundance ~ period, data = sv_ends)
wilcox.test(biomass ~ period, data = sv_ends)
wilcox.test(mean_size ~ period, data = sv_ends)


t.test(abundance ~ period, data = sv_ends)
t.test(biomass ~ period, data = sv_ends)
t.test(mean_size ~ period, data = sv_ends)

ggplot(filter(sv_ends_long), aes(period, value, color = currency)) +
  geom_boxplot() +
  geom_point() +
  facet_wrap(vars(currency), scales = "free", ncol = 4)



ggplot(filter(sv_ends_long, currency == "mean_size"), aes(period, value, color = currency)) +
  geom_boxplot() +
  geom_point() +
  facet_wrap(vars(currency), scales = "free", ncol = 4)
```


### State variable change GAMs

```{r}


library(dplyr)
library(gratia)
library(ggplot2)
load_mgcv()

ts <- read.csv(here::here("analysis", "from_stories", "working_datasets.csv"))

unique_sites <- unique(ts$site_name)

site_dfs <- lapply(unique_sites, FUN = function(site, full_ts) return(filter(full_ts, site_name == site)), full_ts = ts)

source(here::here("analysis", "from_stories", "gam_fxns", "wrapper_fxns.R"))
source(here::here("analysis", "from_stories", "gam_fxns", "sunrise_fxns.R"))

```

```{r just e portal, echo = F}

rats <- filter(ts, site_name == "hartland")

#portal_mean_perc_e <- sum(rats$energy) / sum(rats$abundance)
# 
# rats <- rats %>% 
#   mutate(scaled_energy = energy / portal_mean_perc_e,
#          mean_e = energy / abundance)

rats <- rats %>% 
  mutate(mean_m = biomass / abundance)

rats_long <- rats %>%
  select(year, biomass, abundance, mean_m) %>%
  tidyr::pivot_longer(-year, names_to = "currency", values_to = "value")

ggplot(filter(rats_long), aes(year, value, color = currency)) +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d() +
  facet_wrap(vars(currency), ncol = 1, scales = "free")

```


```{r}
e_mod <- mod_wrapper(rats, response_variable = "biomass", identifier = "site_name", k = 5)
n_mod <- mod_wrapper(rats, response_variable = "abundance", identifier = "site_name", k = 5)
meane_mod <-  mod_wrapper(rats, response_variable = "mean_m", identifier = "site_name", k = 5)

e_samples <- samples_wrapper(e_mod, seed_seed = 1994)
n_samples <- samples_wrapper(n_mod, seed_seed = 1990)
meane_samples <- samples_wrapper(meane_mod, seed_seed = 1977)

joint_samples <- bind_rows(e_samples, n_samples, meane_samples) %>%
  select(year, currency, mean, upper, lower) %>%
  distinct()

ggplot(joint_samples, aes(year, mean, color = currency, fill = currency)) +
  geom_line(aes(year, mean)) +
  geom_ribbon(aes(year, ymin = lower, ymax = upper),  alpha = .25) +
  ggtitle("Fits for energy and abundance") +
  theme_bw() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(vars(currency), scales = "free", ncol = 1)

```


Calculated off the samples, we can get:

* the net change from beginning to end, plus a CI
* a timeseries of the instantaneous rate of change, which we can summarize in order to get a sense of how much backtracking is occurring over the course of the timeseries

## Net change

```{r shortcut to net change}


e_change <- net_change_wrapper(e_samples)
n_change <- net_change_wrapper(n_samples)
meane_change <- net_change_wrapper(meane_samples)

both_change <- bind_rows(e_change, n_change, meane_change)

ggplot(both_change, aes(currency, net_proportional, color = currency)) +
  geom_boxplot() +
  theme_bw() +
  scale_color_viridis_d()

both_summary <- change_summary_wrapper(both_change)

both_summary
```
